Comparison of MCMC algorithms in Stochastic Volatility Models

### 2023-08-21 writeups/research_proposal/propsal.qmd
spacing

---
title: "Comparison of MCMC algorithms in Stochastic Volatility Models"
format:
  pdf:
    fig-pos: H
---

# Background and Motivation
Maybe focus on the MCMC as opposed to SV upfront.

@@

How do we validate the computation and implementation of our sampling strategies and algorithms in increasingly complex models? And how do these form part of the computational modelling workflow?

@@

# Literature Review

## KSC
- KSC (1998) compare the performance of likelihood based inference and Bayesian inference at estimating the fit of these models. 
- Specifically KSC propose a method of sampling using an offset mixture normal approximation with a reweighting procedure to sample from the joint posterior of the SV model. This is estimated using conjugate priors as well as MH.
- The use of simulation and sampling are used as part of model diagnostics and evaluation. This is compared with other estimation strategies (MLE) and volatility models such as autogregressive conditional heteroskedasticity (ARCH).

## Stan HMC + NUTs
There have been many developments in MCMC algorithms since the introduction of classical methods such as GIBBS and MH. Hamiltoninn Monte Carlo (algorithm introduced when, widely available since 2010...?) is a relatively new im.

@@

## Simulation Based Calibration (SBC)
<!-- SBC validates posterior samples and inferences generated by MCMC algorithms. It identifies inconsistencies and bias in computation and estimation of a generative model. This procedure  -->

>>>>

---
title: "Comparison of MCMC algorithms in Stochastic Volatility Models"
format:
  pdf:
    fig-pos: H
    include-in-header:
        text: |
          \usepackage{lipsum}
          \usepackage{setspace}
          \onehalfspacing
          \linespread{1.5}
fontsize: 12pt
geometry: margin=1in
---

<!-- # Background and Motivation
Maybe focus on the MCMC as opposed to SV upfront.

@@

How do we validate the computation and implementation of our sampling strategies and algorithms in increasingly complex models? And how do these form part of the computational modelling workflow?    -->

@@

# Background and Literature Review

Kim Sherpherd and Chib (1998) estimate the discrete time Stochastic Volatility model by proposing a simulation strategy using Bayesian methods. Specifically, KSC propose a method of sampling using an offset mixture normal approximation with a reweighting procedure to sample from the posterior of the SV model parameters. By linearing the model and using a mixture of Gaussian's, they were able to apply a Kalman Filter to estimate the latent volatilities of the state space models. The use of simulation and sampling are used as part of model diagnostics and evaluation. This is compared with other estimation strategies and volatility models such as autogregressive conditional heteroskedasticity (ARCH). Their specificaiton of the Stochastic Volatility model and estimation approaches provide the context for comparing the performance of different MCMC algorithms.

The simulations run by KSC rely on sampling directly from conjugate priors and classic Markov Chain Monte Carlo algorithms, Gibbs and Metropolis Hastings. Since the paper was written, new and more powerful MCMC algorithms have been developed and enabled the estimation of richer and more complicated models. Hamiltonian Monte Carlo, through open source programming languages such as Stan and PyMC, have become

There have been many developments in MCMC algorithms since the introduction of classical methods such as Gibbs and Metropolis Hastings.

Hamiltoninn Monte Carlo (algorithm introduced when, widely available since 2010...?) is a relatively new im.

@@

## 3.3 Simulation Based Calibration (SBC)
<!-- SBC validates posterior samples and inferences generated by MCMC algorithms. It identifies inconsistencies and bias in computation and estimation of a generative model. This procedure  -->
